Labor Market Income and Substitution Effects Explain Contemporary Natural Selection

### 2020-12-18 why-natural-selection.Rmd
Add purifying selection (quadratic) regressions. Rename file.

Since UK Biobank subjects are not representative of the wider population, our
results suggest that naïve estimates of natural selection are at risk of
ascertainment bias. To correct for this, we weight participants using population data. We try three
alternative weighting schemes: weighting by geography, age and
presence/absence of a partner; weighting by age and highest educational
qualification; and age, highest qualification, and age at first live birth, for
women only. Figure \@ref(fig:plot-res-wt) plots selection effects 
among the entire sample, estimated with the three weighting schemes. 
Mean effect sizes across all polygenic scores are
increased by a factor of `r mean_rel$Geographical` (geographical weighting), 
`r mean_rel[["Age/Qualification"]]` (age/qualification) or 
`r mean_rel[["Age/Qual/AFLB"]]` (age/qualification/AFLB). Estimates might be
further affected by weighting on other demographic variables. Since the Biobank
sample seems to be ascertained in ways that shrink estimates of selection
effects, we suspect that our weighted estimates are still conservative.

@@

>>>>

Since UK Biobank subjects are not representative of the wider population, our
results suggest that naïve estimates of natural selection are at risk of
ascertainment bias. To correct for this, we weight participants using population data. We try three
alternative weighting schemes: weighting by geography, age and
presence/absence of a partner; weighting by age and highest educational
qualification; and age, highest qualification, and age at first live birth, for
women only. Figure \@ref(fig:plot-res-wt) plots selection effects 
among the entire sample, estimated with the three weighting schemes. 
Mean effect sizes across all polygenic scores are
increased by a factor of `r mean_rel$Geographical` (geographical weighting), 
`r mean_rel[["Age/Qualification"]]` (age/qualification) or 
`r mean_rel[["Age/Qual/AFLB"]]` (age/qualification/AFLB). Estimates might be
further affected by weighting on other demographic variables. Since the Biobank
sample seems to be ascertained in ways that shrink estimates of selection
effects, we suspect that our weighted estimates are still conservative.[^purifying]

[^purifying]: We also check for purifying selection by estimating (\@ref(eq:regression)) with a quadratic term in $PGS_i$. See Appendix figure \@ref(fig:plot-purifying).

@@

\FloatBarrier
\clearpage
## Purifying and diversifying selection

We rerun equation (\@ref(eq:regression)), adding a quadratic term in $PGS_i$,
and using our age/qualification weights. Scores for hip circumference and drinks
per week show significant purifying selection ($p < 0.05/33$, negative
coefficient). Scores for educational attainment (EA2 and EA3) show significant
diversifying selection ($p < 0.05/33$, positive coefficient), which reduces the
strength of selection against educational attainment at very high levels of the 
PGS. Figure
\@ref(fig:plot-purifying) plots predicted number of children against polygenic
score from these regressions.

```{r plot-purifying, fig.align = "center", fig.cap = "Purifying/diversifying selection: predicted number of children by polygenic score", fig.height = 5}

drake::loadd(res_quadratic)
sig <- res_quadratic %>% 
         filter(grepl("^2", term, fixed = TRUE)) %>% 
         filter(p.value < 0.05/33) %>% 
         pull(score_name)

stopifnot(sig == c("dpw_substance_use", "EA2_noUKB", "EA3_excl_23andMe_UK",
                     "hip_combined"))

sig_all_terms <- res_quadratic %>% filter(score_name %in% sig)

op <- par(mfrow = c(2,2))
for (score in sig) {
  coef <- sig_all_terms %>% filter(score_name == score) %>% pull(estimate)
  curve(
         coef[1] + coef[2] * x + coef[3] * x^2,
         xlim = c(-4, 4), 
         xlab = pretty_names(score),
         ylab = "Predicted N children"
       )
}
par(op)
